The \dealii Library, Version 9.6

### 2024-09-19 9.6/paper.tex
Adjust reference to two previous versions
Improve Tpetra section

\dealii{}'s parallel linear algebra facilities are largely built on
wrappers around functionality provided by the \petsc{} and \trilinos{} libraries.
As a software project, \trilinos{} is designed as a stack of semi-independent
packages building upon each other and working together. 
As a foundation, \teuchos{} offers a set of basic tools,
including parameter handling through XML files or command line interfaces 
and the reference counted pointer \texttt{RCP} which predates 
the smart pointer types introduced in \texttt{C++11}.
The second layer in the stack is a package providing distributed linear algebra 
classes for vectors and (sparse) matrices based on the \petra{}
object model. The third layer is a set of interconnected packages
providing preconditioners, various types of solvers 
as well as more advanced techniques like automatic differentiation, 
time integration and discretization.

Historically, \trilinos{} has implemented \petra{} objects -- such as
parallel vector, sparsity pattern, and sparse matrix classes -- in its
\epetra{} package that uses \MPI{} as its only source of
parallelism. The \dealii{} interfaces to \trilinos{} have therefore traditionally all
been built using \epetra{}.
However, several years ago, \trilinos also introduced the
newer \tpetra{} (``templated'' Petra) package that provides additional shared memory 
parallelism and GPU capabilities by building on \kokkos{}. \tpetra{}
is slated to eventually replace \epetra{}, and the latter is indeed
now deprecated with removal slated for 2025. As a consequence, we will
eventually have to switch all of our \trilinos{} interfaces to
\tpetra{}; the same is true for the need to switch to \tpetra{}-based
sub-packages replacing existing \trilinos sub-packages (for example,
Ifpack2 instead of Ifpack).

In the current release, we have put substantial work into this switch.

\todo[inline]{Sebastian \& Jan Philipp: Please complete}

\begin{itemize}
  \item Short overview capabilities of \epetra (MPI parallel, fixed types) [Thiele]
  \item Short overview capabilities of \tpetra (MPI+X via Kokkos, auto diff types, Templated petra) [Thiele]
    \todo[inline]{Keep discussion of the two items above as short as
      you can -- we're describing the changes in deal.II, not some
      other package. Perhaps roll what you want to say into the
      paragraphs above?}
  \item Current state of the interface (Tpetra based linear algebra, direct solver from Amesos2, preconditioners from Ifpack2) [Thiele \& Kinnewig]
  \item User relevant changes from \epetra: incompatibilities, removed
    options, changes in the constructor. [Thiele \& Kinnewig]
    \todo[inline]{These last two items are the important ones.}
\end{itemize}

@@

The two previous releases of \dealii can be cited as
\cite{dealII92,dealII93}.

>>>>

\dealii{}'s parallel linear algebra facilities (apart from matrix-free capabilities) are largely built on
wrappers around functionality provided by the \petsc{} and \trilinos{} libraries.

Historically, \trilinos{} has implemented distributed linear algebra 
-classes for vectors and (sparse) matrices in its
\epetra{} package that uses \MPI{} as its only source of
parallelism. The \dealii{} interfaces to \trilinos{} have therefore traditionally all
been built using \epetra{}.
However, several years ago, \trilinos also introduced the
newer \tpetra{} (``templated'' Petra) package that provides additional shared memory 
parallelism and GPU capabilities by building on \kokkos{}. \tpetra{}
is slated to eventually replace \epetra{}, and the latter is indeed
now deprecated with removal slated for 2025. As a consequence, we will
eventually have to switch all of our \trilinos{} interfaces to
\tpetra{}; the same is true for the need to switch to \tpetra{}-based
sub-packages replacing existing \trilinos sub-packages (for example,
Ifpack2 instead of Ifpack).

In the current release, we have put substantial work into this switch. The following omits the \texttt{LinearAlgebra} prefix for readability.
\begin{itemize}
  \item The \texttt{TpetraWrappers::Vector} class has been overhauled to also allow specifying
    a memory space, and the \texttt{TpetraWrappers::BlockVector} class has been added.
  \item \texttt{TpetraWrappers::SparseMatrix} and \texttt{TpetraWrappers::SparsityPattern} has been implemented mirroring
    the functionality of the \texttt{TrilinosWrappers::SparseMatrix} \\ and \texttt{TrilinosWrappers::SparsityPattern} classes.
  \item Various \texttt{TpetraWrappers::Precondition*} have been added largely
    mirroring the exising preconditioners in the \texttt{LinearAlgbera::TrilinosWrappers} namespace.
  \item \texttt{TpetraWrappers::SolverDirect} and \texttt{TpetraWrappers::SolverDirectKLU2}
    have been implemented as direct solvers.
\end{itemize}
Most of the interfaces are usable at this point even if some functionality existing in the \texttt{TrilinosWrappers} classes
is still missing, most noticably a wrapper for an algebraic-multigrid preconditioner.

All the \texttt{TpetraWrappers} classes take \texttt{Number} and \texttt{MemorySpace} template arguments
allowing to control the underlying number type and the memory space used for storing the data. Note that the memory space type
also determines whether \tpetra{} takes advantage of GPUs if \texttt{Kokkos} has been configure with a GPU backend.

@@

The two previous releases of \dealii can be cited as
\cite{dealII94,dealII95}.
